true

### 2020-04-17 analysis/robinson2020.Rmd
bumping to r 3.6.3 in dockerfile

---
title: 
    main: Dendrochronological dates confirm a Late Prehistoric population decline in the American Southwest derived from radiocarbon dates
    short: Dendrochronological and radiocarbon dates in the American Southwest
date: "`r format(Sys.Date(), '%B %d, %Y')`"
author:
  - Erick Robinson:
      email: erick.robinson@usu.edu
      institute: [USU]
      correspondence: true
  - Kyle Bocinsky:
      email: kbocinsky@crowcanyon.org
      institute: [CCAC,UM]
      correspondence: false
  - Darcy Bird:
      email: darcy.bird@wsu.edu
      institute: [WSU]
      correspondence: false
  - Jacob Freeman:
      email: jacob.freeman@usu.edu
      institute: [USU]
      correspondence: false
  - Robert L. Kelly:
      email: rlkelly@uwyo.edu 
      institute: [UW]
      correspondence: false
institute:
  - USU: Utah State University
  - CCAC: Crow Canyon Archaeological Center
  - UM: University of Montana
  - WSU: Washington State University
  - UW: University of Wyoming
abstract: >
  The northern American Southwest provides one of the most well-documented cases of human population growth and decline in the world. The geographic extent of this decline in North America is unknown due to the lack of high-resolution palaeodemographic data from regions across and beyond the greater Southwest, where archaeological radiocarbon data is often the only available proxy for investigating these palaeodemographic processes. Radiocarbon time series across and beyond the greater Southwest suggest widespread population collapses from AD 1300--1600. However, radiocarbon data have potential biases caused by variable radiocarbon sample preservation, sample collection, and the non-linearity of the radiocarbon calibration curve. In order to be confident in the wider trends seen in radiocarbon time series across and beyond the greater Southwest, here we focus on regions that have multiple palaeodemographic proxies and compare those proxies to radiocarbon time series. We develop a new method for time series analysis and comparison between dendrochronological data and radiocarbon data. Results confirm a multiple proxy decline in human populations across the Upland US Southwest, Central Mesa Verde, and Northern Rio Grande from AD 1300--1600. These results lend confidence to single proxy radiocarbon-based reconstructions of paleodemography outside the Southwest that suggest post-AD 1300 population declines in many parts of North America.
keywords:
    formatted:
      - dates-as-data
      - radiocarbon summed probability distributions
      - dendrochronology
      - palaeodemography
      - southwestern United States
doi: [DOI HERE]
journal: Preprint
documentclass: sa
bibliography: robinson2020_references.bib
# csl: "philosophical-transactions-of-the-royal-society-b.csl"
csl: "journal-of-archaeological-science.csl"
link-citations: true
output: 
    bookdown::pdf_document2:
      template: templates/template.tex
      latex_engine: xelatex
      keep_tex: true
      fig_caption: yes
      extra_dependencies:
        caption: ["labelfont={bf}"]
      pandoc_args:
      - --lua-filter=templates/scholarly-metadata.lua
      - --lua-filter=templates/author-info-blocks.lua
      - --lua-filter=templates/pagebreak.lua
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  echo = FALSE,
  fig.align = "center",
  fig.path = here::here("analysis/figures/"),
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  dev = "pdf"
)
library(robinson2020)
library(ggplot2)
```

@@

rm(radiocarbon_spd, dendro_spd)
gc(verbose = FALSE)
gc(verbose = FALSE)

>>>>

---
title: 
    main: Dendrochronological dates confirm a Late Prehistoric population decline in the American Southwest derived from radiocarbon dates
    short: Dendrochronological and radiocarbon dates in the American Southwest
date: "`r format(Sys.Date(), '%B %d, %Y')`"
author:
  - Erick Robinson:
      email: erick.robinson@usu.edu
      institute: [USU]
      correspondence: true
  - Kyle Bocinsky:
      email: kbocinsky@crowcanyon.org
      institute: [CCAC,UM]
      correspondence: false
  - Darcy Bird:
      email: darcy.bird@wsu.edu
      institute: [WSU]
      correspondence: false
  - Jacob Freeman:
      email: jacob.freeman@usu.edu
      institute: [USU]
      correspondence: false
  - Robert L. Kelly:
      email: rlkelly@uwyo.edu 
      institute: [UW]
      correspondence: false
institute:
  - USU: Utah State University
  - CCAC: Crow Canyon Archaeological Center
  - UM: University of Montana
  - WSU: Washington State University
  - UW: University of Wyoming
abstract: >
  The northern American Southwest provides one of the most well-documented cases of human population growth and decline in the world. The geographic extent of this decline in North America is unknown due to the lack of high-resolution palaeodemographic data from regions across and beyond the greater Southwest, where archaeological radiocarbon data is often the only available proxy for investigating these palaeodemographic processes. Radiocarbon time series across and beyond the greater Southwest suggest widespread population collapses from AD 1300--1600. However, radiocarbon data have potential biases caused by variable radiocarbon sample preservation, sample collection, and the non-linearity of the radiocarbon calibration curve. In order to be confident in the wider trends seen in radiocarbon time series across and beyond the greater Southwest, here we focus on regions that have multiple palaeodemographic proxies and compare those proxies to radiocarbon time series. We develop a new method for time series analysis and comparison between dendrochronological data and radiocarbon data. Results confirm a multiple proxy decline in human populations across the Upland US Southwest, Central Mesa Verde, and Northern Rio Grande from AD 1300--1600. These results lend confidence to single proxy radiocarbon-based reconstructions of paleodemography outside the Southwest that suggest post-AD 1300 population declines in many parts of North America.
keywords:
    formatted:
      - dates-as-data
      - radiocarbon summed probability distributions
      - dendrochronology
      - palaeodemography
      - southwestern United States
doi: [DOI HERE]
journal: Preprint
# documentclass: sa
bibliography: robinson2020_references.bib
# csl: "philosophical-transactions-of-the-royal-society-b.csl"
csl: "journal-of-archaeological-science.csl"
link-citations: true
output: 
    bookdown::pdf_document2:
      # template: templates/template.tex
      latex_engine: xelatex
      keep_tex: true
      fig_caption: yes
      toc: false
      extra_dependencies:
        caption: ["labelfont={bf}"]
      pandoc_args:
      - --lua-filter=templates/scholarly-metadata.lua
      - --lua-filter=templates/author-info-blocks.lua
      # - --lua-filter=templates/pagebreak.lua
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  echo = FALSE,
  fig.align = "center",
  fig.path = here::here("analysis/figures/"),
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  dev = "pdf"
)
library(robinson2020)
library(ggplot2)
```

@@

### 2020-04-17 analysis/robinson2020.tex
bumping to r 3.6.3 in dockerfile

\documentclass[
]{sa}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\author{
Erick Robinson\textsuperscript{1,*}

\And
Kyle Bocinsky\textsuperscript{2,3}

\And
Darcy Bird\textsuperscript{4}

\And
Jacob Freeman\textsuperscript{1}

\And
Robert L. Kelly\textsuperscript{5}

\title{Dendrochronological dates confirm a Late Prehistoric population decline in the American Southwest derived from radiocarbon dates}

\Plainauthor{Erick Robinson\textsuperscript{1,*}, Kyle Bocinsky\textsuperscript{2,3}, Darcy Bird\textsuperscript{4}, Jacob Freeman\textsuperscript{1}, Robert L. Kelly\textsuperscript{5}}

\Shorttitle{Dendrochronological and radiocarbon dates in the American Southwest}

\Abstract{
The northern American Southwest provides one of the most well-documented cases of human population growth and decline in the world. The geographic extent of this decline in North America is unknown due to the lack of high-resolution palaeodemographic data from regions across and beyond the greater Southwest, where archaeological radiocarbon data is often the only available proxy for investigating these palaeodemographic processes. Radiocarbon time series across and beyond the greater Southwest suggest widespread population collapses from AD 1300--1600. However, radiocarbon data have potential biases caused by variable radiocarbon sample preservation, sample collection, and the non-linearity of the radiocarbon calibration curve. In order to be confident in the wider trends seen in radiocarbon time series across and beyond the greater Southwest, here we focus on regions that have multiple palaeodemographic proxies and compare those proxies to radiocarbon time series. We develop a new method for time series analysis and comparison between dendrochronological data and radiocarbon data. Results confirm a multiple proxy decline in human populations across the Upland US Southwest, Central Mesa Verde, and Northern Rio Grande from AD 1300--1600. These results lend confidence to single proxy radiocarbon-based reconstructions of paleodemography outside the Southwest that suggest post-AD 1300 population declines in many parts of North America.
}

\Keywords{dates-as-data, radiocarbon summed probability distributions, dendrochronology, palaeodemography, southwestern United States}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
          }

%% Publication information
%% Edit as needed depending on status of manuscript.
%% Replace Preprint with journal short name of accepted
%% Pubs if copyright is open compliant. Add punctuation
%% for formatting as desired.
\Journal{Preprint}
% \Volume{}
% \Issue{} %Enter as "(Issue)" for APA
% \Pages{} %Enter as "pp. m---n" for APA
% \Year{}
\Submitdate{\today} %Enter date you submitted preprint for review
\Acceptdate{\today} %Enter date your paper was accepted for publication
\DOI{
DOI HERE
} %Enter final DOI here

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

% Pandoc header
\usepackage[labelfont={bf}]{caption}

\begin{document}

@@

\begin{verbatim}
#>           used (Mb) gc trigger   (Mb) limit (Mb)  max used   (Mb)
#> Ncells  716943 38.3    1232191   65.9         NA   1232191   65.9
#> Vcells 1639461 12.6  508769831 3881.7     716800 507690663 3873.4
#>           used (Mb) gc trigger   (Mb) limit (Mb)  max used   (Mb)
#> Ncells  717083 38.3    1232191   65.9         NA   1232191   65.9
#> Vcells 1639786 12.6  407015865 3105.3     716800 507690663 3873.4
\end{verbatim}

@@

This report is part of a research compendium developed using the \texttt{rrtools} package for reproducible research (Marwick, \protect\hyperlink{ref-rrtools}{2019}, \protect\hyperlink{ref-Marwick2017}{2017}). This report was generated on 2020-04-17 08:31:33 using the following computational environment and dependencies:

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdfauthor={Erick Robinson1,; Kyle Bocinsky2,3; Darcy Bird4; Jacob Freeman1; Robert L. Kelly5},
  pdfkeywords={true},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage[labelfont={bf}]{caption}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

\title{true}
\author{Erick Robinson\textsuperscript{1,*} \and Kyle Bocinsky\textsuperscript{2,3} \and Darcy Bird\textsuperscript{4} \and Jacob Freeman\textsuperscript{1} \and Robert L. Kelly\textsuperscript{5}}
\date{April 17, 2020}

\begin{document}
\maketitle
\begin{abstract}
The northern American Southwest provides one of the most well-documented cases of human population growth and decline in the world. The geographic extent of this decline in North America is unknown due to the lack of high-resolution palaeodemographic data from regions across and beyond the greater Southwest, where archaeological radiocarbon data is often the only available proxy for investigating these palaeodemographic processes. Radiocarbon time series across and beyond the greater Southwest suggest widespread population collapses from AD 1300--1600. However, radiocarbon data have potential biases caused by variable radiocarbon sample preservation, sample collection, and the non-linearity of the radiocarbon calibration curve. In order to be confident in the wider trends seen in radiocarbon time series across and beyond the greater Southwest, here we focus on regions that have multiple palaeodemographic proxies and compare those proxies to radiocarbon time series. We develop a new method for time series analysis and comparison between dendrochronological data and radiocarbon data. Results confirm a multiple proxy decline in human populations across the Upland US Southwest, Central Mesa Verde, and Northern Rio Grande from AD 1300--1600. These results lend confidence to single proxy radiocarbon-based reconstructions of paleodemography outside the Southwest that suggest post-AD 1300 population declines in many parts of North America.
\end{abstract}

@@

@@

This report is part of a research compendium developed using the \texttt{rrtools} package for reproducible research (Marwick, \protect\hyperlink{ref-rrtools}{2019}, \protect\hyperlink{ref-Marwick2017}{2017}). This report was generated on 2020-04-17 11:04:15 using the following computational environment and dependencies:
